Open archaeology, open source?

### 2023-09-20 analysis/_03-data_methodology.qmd
Note TODO
Expand network analysis

@@

>>>>

<!-- TODO:
* Introduce git, GitHub, etc. and describe briefly how they work for non-users
* Define software jargon we use (repo, commit, etc.) – maybe as a glossary table?
-->

@@

```{r data-oarch_contributors}
oarch |>
  unnest(gh_contribs) |>
  drop_na(contributor) |>
  group_by(contributor) |>
  mutate(
    repo_owner = if_else(str_extract(repo, "[^/]+") == contributor, "own", "other")
  ) |>
  group_by(contributor, repo_owner) |>
  summarise(repos = n(), contribs = sum(contributions)) |>
  pivot_wider(c(contributor), names_from = repo_owner, 
              values_from = c(repos, contribs), values_fill = 0) |>
  mutate(
    repos = repos_other + repos_own,
    contribs = contribs_other + contribs_own,
    .after = contributor
  ) ->
  oarch_contributors
```

### 2023-09-20 analysis/_06-network.qmd
Expand network analysis

```{r graph-data, eval=FALSE}
# TODO: refactor some of this as functions?
oarch |>
  drop_na(gh_repo) |>
  unite(tags, c(tag1, tag2, tag3, tag4, tag5)) |>
  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
  oarch_graph

@@

```{r plot-graph-repo-user, eval=FALSE}
oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
                                   edges_user_to_repo(oarch_graph))

g_repo_user <- oarch_graph_repo_user |>
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
  geom_node_point(aes(colour = node_type)) +
  scale_edge_alpha(trans = "log")

@@

```{r plot-graph-repo-repo, eval=FALSE}
oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category),
                                   edges_common_users(oarch_graph),
                                   directed = FALSE)

oarch_graph_repo_repo |>
  ggraph(layout = "graphopt") +
  geom_edge_link(aes(alpha = n)) +
  geom_node_point(aes(colour = category)) +
  scale_edge_alpha(trans = "log") +
  labs(edge_alpha = "Common contributors")
```

```{r graph-user-user, eval=FALSE}
oarch_graph_user_user <- tbl_graph(
  nodes = nodes_users(oarch_graph), 
  edges = edges_common_repos(oarch_graph),
  directed = FALSE
)
```
```{r plot-graph-user-user, eval=FALSE}
oarch_graph_user_user |>
  mutate(betweenness = centrality_betweenness(weights = n)) ->
  oarch_graph_user_user

oarch_graph_user_user |> 
  ggraph(layout = "graphopt") +
  geom_edge_link(aes(alpha = n)) +
  # geom_node_point() +
  scale_edge_alpha(trans = "log") +
  labs(edge_alpha = "Common repositories")
```

@@

```{r plot-graph-repo-repo-edge-betweenness, eval=FALSE}
oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
                                   edges_common_users(oarch_graph),
                                   directed = FALSE)

iplot <- oarch_graph_repo_repo |>
  ggraph(layout = "graphopt") +
  geom_edge_link(aes(alpha = n)) +
  #geom_node_point(aes(colour = group)) +
  #geom_node_text(aes(label = group)) +
  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
  scale_edge_alpha(trans = "log") +
  labs(edge_alpha = "Common contributors")

@@

```{r plot-graph-repo-repo-edge-betweenness-dendrogram, eval=FALSE}
oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
                                   edges_common_users(oarch_graph),
                                   directed = TRUE)

@@

**NOTE:** Create a table containing the nodes with highest betweenness centrality from the repo-repo graph. Columns should specify the name of the repository, the number of commits, the category, and tags. If possible also include a summary of the pulse as a line chart with x axis as time and y axis as activity. This will allow us to identify the characteristics of these repositories that prompt greater integration.

@@

**NOTE:** Plot a graph that renders nodes' sizes based on their betweenness centrality.

```{r summary-graph-user-user, eval=FALSE}
oarch_graph_user_user |>
  mutate(degree = centrality_degree(), 
         betweenness = centrality_betweenness(weights = n), 
         strength = centrality_degree(weights = n)) |>
  filter(degree > 0) |> 
  data.frame()
```

@@

@@

**NOTE:** I have no idea what these next two charts depict.

```{r plot-user-betweenness-dist, eval=FALSE}
oarch_graph_user_user |>
  mutate(
    betweenness = centrality_betweenness(weights = n)
  ) |>
  data.frame() |>
  ggplot(aes(x = betweenness)) +
  geom_density(bw = "nrd") +
  scale_x_log10(labels = scales::number)
```

```{r plot-user-betweenness-activity, eval=FALSE}
oarch_graph_user_user |>
  mutate(
    degree = centrality_degree(), 
    betweenness = centrality_betweenness(weights = n), 
    strength = centrality_degree(weights = n)
  ) |>
  as_tibble() |> 
  left_join(oarch_contributors, by = c("user" = "contributor")) |>
  ggplot(aes(betweenness, contribs)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "lm")
```

>>>>

```{r data-graph}
# TODO: refactor some of this as functions?
oarch |>
  drop_na(gh_repo) |>
  unite(tags, c(tag1, tag2, tag3, tag4, tag5)) |>
  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
  oarch_graph

@@

```{r extract-graph-repo-user, eval=FALSE}
oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
                                   edges_user_to_repo(oarch_graph))
```

```{r fig-graph-repo-user, eval=FALSE}
g_repo_user <- oarch_graph_repo_user |>
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
  geom_node_point(aes(colour = node_type)) +
  scale_edge_alpha(trans = "log")

@@

@@

<!-- TODO: Note that we remove disconnected repos? (and how many?) -->

```{r extract-graph-repo-repo}
oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
                                   edges_common_users(oarch_graph),
                                   directed = FALSE)

oarch_graph_repo_repo |>
  mutate(degree = centrality_degree()) |>
  filter(degree > 1) |>
  mutate(group = as.character(group_edge_betweenness())) ->
  oarch_graph_repo_repo
```

```{r fig-graph-repo-repo, eval=FALSE}
oarch_graph_repo_repo |>
  ggraph(layout = "graphopt") +
  geom_edge_link(aes(alpha = n)) +
  geom_node_point(aes(colour = category)) +
  scale_edge_alpha(trans = "log") +
  labs(edge_alpha = "Common contributors")
```

```{r fig-graph-repo-repo, eval=FALSE}
iplot <- oarch_graph_repo_repo |>
  ggraph(layout = "graphopt") +
  geom_edge_link(aes(alpha = n)) +
  #geom_node_point(aes(colour = group)) +
  #geom_node_text(aes(label = group)) +
  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
  scale_edge_alpha(trans = "log") +
  labs(edge_alpha = "Common contributors")

@@

```{r fig-graph-repo-repo-edge-betweenness-dendrogram, eval=FALSE}
oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
                                   edges_common_users(oarch_graph),
                                   directed = TRUE)

@@

```{r tbl-repo-centrality}
#| tbl-cap: Most central repositories in the repository–repository network
oarch_graph_repo_repo |>
  mutate(centrality = centrality_betweenness()) |>
  as_tibble() |>
  arrange(-centrality) |>
  slice(1:10) |>
  left_join(oarch, c("repo" = "gh_repo")) |>
  mutate(n_commits = map_int(gh_commits, nrow)) |>
  unite(tags, tag1:tag5, sep = "; ", na.rm = TRUE) |>
  select(repo, centrality, category, tags, n_commits) |>
  gt() |>
  cols_label(
    repo = "Repository", 
    centrality = "Betweenness Centrality",
    category = "Category",
    tags = "Tags",
    n_commits = "Commits"
  )
```

<!-- TODO: What does tbl-repo-centrality tell us? Can we interrogate it further with a graph? -->

@@

<!-- TODO: Note that we remove disconnected users? (and how many?) -->

```{r graph-user-user}
oarch_graph_user_user <- tbl_graph(
  nodes = nodes_users(oarch_graph), 
  edges = edges_common_repos(oarch_graph),
  directed = FALSE
)

oarch_graph_user_user |>
  mutate(degree = centrality_degree()) |>
  filter(degree > 1) |>
  mutate(
    centrality = centrality_betweenness(weights = n),
    group = as.character(group_edge_betweenness())
  ) ->
  oarch_graph_user_user
```

<!--
 NOTE, ZACK: the graph below is what you requested (nodes' sizes based on their betweenness centrality), but I'm unsure we will ever be able to get these graph plots legible enough to actually include. – JR -->

```{r fig-graph-user-user-centrality, eval=TRUE}
oarch_graph_user_user |> 
  ggraph(layout = "graphopt") +
  # geom_edge_link(aes(alpha = n)) +
  geom_edge_arc(alpha = 0.05) +
  geom_node_point(aes(size = centrality), 
                  shape = 21, colour = "#000000", fill = "#ffffff") +
  # scale_edge_alpha(trans = "log") +
  # labs(edge_alpha = "Common repositories") +
  theme_graph()
```

@@

<!-- JR: I'm not sure about this. Either practically (can we reliably determine all this from public info?) or ethically (not necessarily that it is unethical; but we haven't really considered the implications of analysing named individuals in this study) -->

```{r tbl-user-centrality}
#| tbl-cap: Most central users in the user–user network
oarch_graph_user_user |>
  mutate(centrality = centrality_betweenness(weights = n)) |>
  as_tibble() |> 
  arrange(-centrality) |>
  slice(1:10) |>
  left_join(oarch_contributors, by = c("user" = "contributor")) |>
  select(user, centrality, repos, contribs_own, contribs_other) |>
  gt() |>
  cols_label(
    user = "User",
    centrality = "Centrality",
    contribs_own = "Own repositories",
    contribs_other = "Others' repositories"
  ) |>
  tab_spanner("Contributions", c(contribs_own, contribs_other))
```

@@

```{r fig-user-betweenness, eval=FALSE}
#| fig-cap: Distribution of betweenness centrality values in the user–user network
oarch_graph_user_user |>
  mutate(
    betweenness = centrality_betweenness(weights = n)
  ) |>
  data.frame() |>
  ggplot(aes(x = betweenness)) +
  geom_density(bw = "nrd") +
  scale_x_log10(labels = scales::number)
```

```{r fig-user-betweenness-activity, eval=FALSE}
#| fig-cap: Correlation between user's centrality (betweenness) in the network and their total number of contributions
oarch_graph_user_user |>
  mutate(
    degree = centrality_degree(), 
    betweenness = centrality_betweenness(weights = n), 
    strength = centrality_degree(weights = n)
  ) |>
  as_tibble() |> 
  left_join(oarch_contributors, by = c("user" = "contributor")) |>
  ggplot(aes(betweenness, contribs)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_smooth(method = "lm")
```
